%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Acknowledgments}{acknowledgments}
\phantomsection
\manualmark
\markboth{\spacedlowsmallcaps{Acknowledgments}}{\spacedlowsmallcaps{Acknowledgments}}%
\addtocontents{toc}{}%
\addcontentsline{toc}{chapter}{\tocEntry{Acknowledgments}}%
\chapter*{Acknowledgments}%
% \bigskip

%to include: \emph{Davis}\\ 
% This thesis would not be possible without the help and support from colleagues, friends and family. 

I have many fond memories working at the theory of condensed matter department, a place that often felt like a second home to me. Playing music on Fridays, the Seinfeld video of the day. Secretly keeping track of who entered our office and performing data analysis on it with \emph{Guus} (\emph{Lennert} won by most visits). What started as the daily coffee break with \emph{Matteo} and \emph{Alessandro}, was taken over by \emph{Achille} and \emph{Andrei}. Walking to work with \emph{Josien}. The daily zig-zag walks with \emph{Sanne}. Sharing weekend stories with \emph{Belinda}. Frequent discussions with \emph{Matthieu} about Magic: The Gathering, politics, economy and any other topic. Complaining with \emph{Taha} that our code never seems to work.
%Akwardly waving at \emph{Marion} every day. Discussing inappropriate things with \emph{Marion}. Acting crazy together with \emph{Marion} ("beeeuh").
The vodka parties at \emph{Jaap}'s place. \emph{Marion} drunkenly shouting Russian curse words at \emph{Brenner} in front of his daughter and wife. The times we played counter strike on Friday afternoon. The time \emph{Andre Bagrov} complained that the menu at caf\'e de Plak had only two meat options. 
%The time \emph{Lennert} \emph{Guus} and \emph{I} did a Monkey Island speed run and I lost miserably because I couldn't remember how to get the Jawbreaker from Cuthroat Bill.
Writing and sharing RefterPy so you can check the Refter menu from the terminal. The time \emph{Tom} accidentally used up all the computation time at Cartesius without saving any of his data. \emph{Bektur} exercising in the office to keep his focus up. Dancing with \emph{Askar}. Regretting hiking Glacier 3000 in Diablerets, Switzerland together with \emph{Misha Baglai} wearing only flip-flops. The time \emph{Zhenya} invited me to spend winter holidays in Moscow at -27 degrees. The time when \emph{Roel} and \emph{Ragnar} barged in my office singing happy birthday. After-work burgers with \emph{Ruben} and \emph{Matthieu} at the HAN student canteen. \emph{Majka} giving me surprise hugs. Walking through park Brakkenstein with \emph{David} discussing life and complaining about academia. Random coffee breaks with \emph{Thom}, lunch and dinners with \emph{Marion}. Casual visits by \emph{Bram}. %Discussing with \emph{Guus} how we could show each other something on our screens without standing up.
Discussing data privacy and many other topics with \emph{Arthur}. \emph{Annalisa}'s tiramisu. The yearly Sinterklaas celebration at TCM, in particular the time when one of the \emph{Sasha}s ate one of the presents, leaving $N-1$ gifts for $N$ people. The time another \emph{Sasha} asked if it is illegal to call the police after he was told to stop reporting the woman with a beard he saw on the street. \emph{Misha Katsnelson}'s stories about Soviet times during the coffee breaks and lunches. The countless stories of \emph{Misha Titov} involving near-death experiences, Russian aluminum mafia and children-gangs of Krasnoyarsk. The awkward silences during some of the weekly department cake breaks. The time \emph{Bob} saved \emph{Sanne} and \emph{I} with his banana we could use to bake cookies in the department's kitchen. All the brownies we ate. Not being able to graduate my Master's in time, because I forgot to hand in a report for a course of \emph{Shengjun}.
% All the times \emph{I} baked brownies and apple pies.
The time \emph{Ivan} retrieved his stolen hard drive from the Camorra. All the times when \emph{Koen} effortlessly shows his communication skills in 4.5 different languages. Playing jianzi with \emph{Edo}, \emph{Peilang}, \emph{Jin}, \emph{Guodong} and \emph{Amanda}. Discussing TikZ with \emph{Erik}. 
%Random games of chess with \emph{Peilang}, \emph{Misha Titov}, and \emph{Askar}.
Dressing up as pikachu to make a PhD video for \emph{Remko}. \emph{Hylke} being the first to join for Friday afternoon drinks. The time all participants of a summer school in Moscow ended up with a week long diarrhea after food poisoning. Sharing craft-beer enthusiasm with \emph{Dima}. Ice skating with \emph{Mariana}, tennis with \emph{Lisa}. The time \emph{Manuel} took over my apartment. The time I randomly met \emph{Michel}, \emph{Clement}'s best friend, in a bar in Leiden.  
% The time we built a five meter snow-penis in front of \emph{Zhenya}'s office.
%The time \emph{Lennert}, \emph{Erik}, \emph{Edo} and \emph{I} took an afternoon off to play a 3 hour game of Shogun.

Outside of work, many good times could be found as well. The time \emph{Andre}, \emph{Matteo} and \emph{I} surprised \emph{Samy} in Toulouse. Sunday evening board game nights. The time \emph{Dave}, \emph{Cas}, \emph{Clarisse} and \emph{Jos} pushed me to join a cooking show when I was hungover. The times that \emph{M\'elanie} shared stories from the hospital's urology department she worked at. Going to TEDx-Maastricht with \emph{Pam}. Meeting up with \emph{Alex}, my grandfather's best friend's grandson. Celebrating \emph{Bj\"orn}'s birthday party at his summer house in \"Oland, Sweden. 
Creating ugly Christmas card photos with \emph{Bram} and \emph{Thom}. The time I hiked up the Sanctuary of the Madonna di San Luca with \emph{Lennart}. 
%All the times playing Magic: The Gathering in Arnhem with \emph{Ruben}, \emph{Matthieu} and in Nijmegen with \emph{Joost} and \emph{Derrick}.
Brewing beer with the \emph{brouwcie} in front of the university sports center, with \emph{Nick} and \emph{Edo} in Utrecht, or on \emph{Ruben}'s balcony with \emph{Ruben} and \emph{Matthieu}. The time we went indoor-skiing with \emph{Wen}, \emph{Edo} and \emph{Marion}. Playing badminton with \emph{Kate} in Kleve. The times \emph{Fabries} shows how easy it is to beat me at Hearthstone. Partying with \emph{Laura} and \emph{Marcella}. Going to the gym with \emph{Robert} and \emph{Jop}. Playing a horror survival game with the lights out with \emph{Kiki}. The time we went to the casino with \emph{Dave}, \emph{Yolanda} and \emph{Jesus} during the vierdaagsefeesten.  Visiting \emph{Judith} in Essen. Going to the down the rabbit hole-festival every other year and many other activities with \emph{Jop}, \emph{Thom}, \emph{Bram}, \emph{Edo}, \emph{Bob}, \emph{Majka}, \emph{Sanne}, \emph{Josien}, \emph{Arthur}, \emph{Judith}, \emph{Kiane}, \emph{Maren}, \emph{Sam}, \emph{Sander}, \emph{Vera}, \emph{Yvonne}, \emph{Belle} \& \emph{Guido}. \\[0.5em]

Of course, a special thanks to my \emph{parents}. Faworki and lemon cake on my birthdays, dinners every weekend. The countless times I receive good advice. The big influence my brother, \emph{Maciek}, had on me while growing up. The times you taught me how to build a calculator in php and how to use regular expressions when I was just 13 years old.  \\[0.5em]

And of course, spending the last two months of my PhD in lockdown with my sweet and lovely girlfriend, \emph{Lan}$\heartsuit$.

\begin{flushright} - \emph{Robert Sokolewicz, October 2020}\end{flushright}
